Calculate average distance to ports from known mariculture locations

### 2022-01-28 marine/salmon/salmon_farms/STEP2_salmon_farm_locations.Rmd
made the figures!
updated for resubmission!

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"
fao_salmon <- read_csv("marine/salmon/salmon_farm/data/fao_mariculture_salmon.csv") %>%
     dplyr::filter(year == 2017)

@@

Source: 
http://aquaculture.scotland.gov.uk/
```{r}
## get scotland data: http://aquaculture.scotland.gov.uk/
# Load in data and select sites that have been producing in the last 3 years

#### TODO #### This could probably be improved by setting up a way to webscrape these data... not sure how this was compiled previously.

scotland_csv <- as.data.frame(read_csv(file.path(path, "Scotland/Scotland_site_details.csv")))

# Subset to salmon farms
sub_scotland <- scotland_csv %>% 
  dplyr::select(ID = `Marine Scotland Site ID`, Easting, Northing, `Producing in Last 3 Years`, `Aquaculture Type`, `Water Type`, Species, subregion = `Local Authority` ) %>% 
  filter(`Producing in Last 3 Years` == c("Yes", "New"),
         `Water Type` %in% c("Seawater", "Freshwater and Seawater"), 
         grepl("Salmon", Species))

@@

write_csv(farms_scotland, here("marine/salmon/salmon_farm/data/known_farm_sites_2_4_5/farms_scotland.csv"))

@@

farms_ps <- read.csv(file.path(path, "USA/PugetSound/USA_PugetSound_AtSalmon.csv")) %>%
  mutate(iso3c = "USA") %>%
  dplyr::select(lon, lat, iso3c) %>%
  mutate(data_quality = 2)

write_csv(farms_ps, here("marine/salmon/salmon_farm/data/known_farm_sites_2_4_5/farms_ps_usa.csv"))

@@

## US Northeast
Source:
Northeast Regional Ocean Council, 2018. Northeast ocean data: aquaculture. [Accessed January 17, 2020] https://www.northeastoceandata.org/data-download/

@@

us_ne <- st_read(file.path(path,"USA/NE/US_NE_Aquaculture.shp"))

# id salmon and active sites
salmon_species <- grep("Salmon|salmon", unique(us_ne$allSpecies), value=TRUE)

us_ne_salmon <- us_ne %>%
  filter(allSpecies %in% salmon_species) %>%
  filter(status == "Active")

us_ne_salmon_mol <- st_transform(us_ne_salmon, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
us_ne_salmon_pts <- st_centroid(us_ne_salmon_mol)
# check
plot(st_geometry(us_ne_salmon_mol))
plot(st_geometry(st_centroid(us_ne_salmon_mol)), add=TRUE)

us_ne_salmon_pts_latlon <- st_transform(us_ne_salmon_pts, crs = "+init=epsg:4326")

farms_us_ne <- data.frame(do.call(rbind, st_geometry(us_ne_salmon_pts_latlon))) 
names(farms_us_ne) <- c("lon", "lat")
farms_us_ne <- farms_us_ne %>%
  mutate(iso3c = "USA") %>%
  mutate(data_quality = 2)

write_csv(farms_us_ne, here("marine/salmon/salmon_farm/data/known_farm_sites_2_4_5/farms_ne_usa.csv"))

@@

@@

farm_files <- list.files(here("marine/salmon/salmon_farm/data/known_farm_sites_2_4_5"), full=TRUE)
all_farms = plyr::ldply(farm_files, read_csv)

@@

write.csv(all_farms_1, here("marine/salmon/salmon_farm/data/global_salmon_farm_lat_lon.csv"), row.names = FALSE)

write.csv(all_farms, here("marine/salmon/salmon_farm/data/global_salmon_farm_lat_lon_quality.csv"), row.names = FALSE)

>>>>

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"
fao_salmon <- read_csv("marine/salmon/salmon_farms/data/fao_mariculture_salmon.csv") %>%
     dplyr::filter(year == 2017)

@@

Source: 
http://aquaculture.scotland.gov.uk/

To download:

http://aquaculture.scotland.gov.uk/data/site_details.aspx

Click "CSV export"

## load in data and select active sites, "Seawater" and "Seawater and Freshwater sites", and any sites that contain "Salmon" "Arctic Char" or "Rainbow Trout"

scotland_csv <- as.data.frame(read_csv(file.path(path, "Scotland/Scotland_site_details_2021_raw.csv")))

unique(scotland_csv$Species)

# Subset to salmonidae farms
sub_scotland <- scotland_csv %>% 
  dplyr::select(ID = `Marine Scotland Site ID`, Easting, Northing, `Producing in Last 3 Years`, `Aquaculture Type`, `Water Type`, Species, subregion = `Local Authority` ) %>% 
  filter(`Producing in Last 3 Years` %in% c("Yes", "New"),
         `Water Type` %in% c("Seawater", "Freshwater and Seawater"),
         str_detect(Species, "Salmon|Trout"))

unique(sub_scotland$Species)

## While there are >350 sites in the data which are deemed to have produced Salmonidae species, only 110 of them have operated in the past 3 years, which is what we include in our map.

@@

write_csv(farms_scotland, here("marine/salmon/salmon_farms/data/known_farm_sites_2_4_5/farms_scotland.csv"))

@@

farms_ps <- read.csv(file.path(path, "USA/PugetSound/USA_PugetSound_AtSalmon.csv")) %>%
  mutate(iso3c = "USA") %>%
  dplyr::select(lon, lat, iso3c) %>%
  mutate(data_quality = 2) %>%
  mutate(subregion = "Washington")

write_csv(farms_ps, here("marine/salmon/salmon_farms/data/known_farm_sites_2_4_5/farms_ps_usa.csv"))

@@

## US Northeast
DEPRECATED:
Source:
Northeast Regional Ocean Council, 2018. Northeast ocean data: aquaculture. [Accessed January 17, 2020]

@@

# us_ne <- st_read(file.path(path,"USA/NE/US_NE_Aquaculture.shp"))
# 
# 
# # id salmon and active sites
# salmon_species <- grep("Salmon|salmon", unique(us_ne$allSpecies), value=TRUE)
# 
# us_ne_salmon <- us_ne %>%
#   filter(allSpecies %in% salmon_species) %>%
#   filter(status == "Active")
# 
# us_ne_salmon_mol <- st_transform(us_ne_salmon, crs = "+init=epsg:3035") # convert to equal area projection to find centroids.
# us_ne_salmon_pts <- st_centroid(us_ne_salmon_mol)
# # check
# plot(st_geometry(us_ne_salmon_mol))
# plot(st_geometry(st_centroid(us_ne_salmon_mol)), add=TRUE)
# 
# us_ne_salmon_pts_latlon <- st_transform(us_ne_salmon_pts, crs = "+init=epsg:4326")
# 
# farms_us_ne <- data.frame(do.call(rbind, st_geometry(us_ne_salmon_pts_latlon)))
# names(farms_us_ne) <- c("lon", "lat")
# farms_us_ne <- farms_us_ne %>%
#   mutate(iso3c = "USA") %>%
#   mutate(data_quality = 2)
# 
# write_csv(farms_us_ne, here("marine/salmon/salmon_farm/data/known_farm_sites_2_4_5/farms_ne_usa.csv"))

@@

# United States

## Maine

Locations of aquaculture leases in Maine, 2019. Maine Department of Marine Resources. https://dmr-maine.opendata.arcgis.com/datasets/maine::mainedmr-aquaculture-aq-leases/about

```{r}
maine_aq_lease <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_-_AQ_Leases/MaineDMR_Aquaculture_-_AQ_Leases.shp"))

# mapview(maine_aq_lease)

maine_aq_lease_active <- maine_aq_lease %>%
  dplyr::filter(STATUS == "A") %>% # 149 leases
  dplyr::filter(str_detect(PRIMARYSP, "Salmon")) # 22 leases.. matches the northeast data

us_me_salmon_mol <- st_transform(maine_aq_lease_active, crs = "+init=epsg:3035") # convert to equal area projection to find centroids.
us_me_salmon_pts <- st_centroid(us_me_salmon_mol)
# check
plot(st_geometry(us_me_salmon_mol))
plot(st_geometry(st_centroid(us_me_salmon_mol)), add=TRUE)

us_me_salmon_pts_latlon <- st_transform(us_me_salmon_pts, crs = "+init=epsg:4326")

farms_us_me <- data.frame(do.call(rbind, st_geometry(us_me_salmon_pts_latlon)))
names(farms_us_me) <- c("lon", "lat")
farms_us_me <- farms_us_me %>%
  mutate(iso3c = "USA") %>%
  mutate(data_quality = 2) %>%
  mutate(subregion = "Maine")

write_csv(farms_us_me, here("marine/salmon/salmon_farms/data/known_farm_sites_2_4_5/farms_maine_usa.csv"))

@@

farm_files <- list.files(here("marine/salmon/salmon_farms/data/known_farm_sites_2_4_5"), full=TRUE)
all_farms = plyr::ldply(farm_files, read_csv)

@@

write.csv(all_farms_1, here("marine/salmon/salmon_farms/data/global_salmon_farm_lat_lon.csv"), row.names = FALSE)

write.csv(all_farms, here("marine/salmon/salmon_farms/data/global_salmon_farm_lat_lon_quality.csv"), row.names = FALSE)

### 2022-01-28 marine/salmon/salmon_farms/STEP3_gapfill_salmon.Rmd
updated for resubmission!

write.csv(tonnes_per_farm, "output/salmon_comp_tonnes_per_farm.csv", row.names = FALSE)

avg_global_salmon_tonnes_per_farm = mean(tonnes_per_farm$tonnes_per_farm)
avg_global_salmon_tonnes_per_farm # 927.156

med_global_salmon_tonnes_per_farm = median(tonnes_per_farm$tonnes_per_farm)
med_global_salmon_tonnes_per_farm # 520.5484
```

@@

write.csv(number_additional_points, here("marine/salmon/salmon_farm/output/salmon_points_to_add_final.csv"), row.names = FALSE)
```

@@

write.csv(all_known_farms_final, here("marine/salmon/salmon_farm/output/all_known_salmon_farms_final.csv"), row.names = FALSE)
```

>>>>

write.csv(tonnes_per_farm, "marine/salmon/salmon_farms/output/salmon_comp_tonnes_per_farm.csv", row.names = FALSE)

avg_global_salmon_tonnes_per_farm = mean(tonnes_per_farm$tonnes_per_farm)
avg_global_salmon_tonnes_per_farm # 927.156 v1; 938.2249 v2 ; 866.4024 v3 after scotland correction

med_global_salmon_tonnes_per_farm = median(tonnes_per_farm$tonnes_per_farm)
med_global_salmon_tonnes_per_farm # 520.5484 v1; 520.5484 v2 ; 520.5484 v3
```

@@

write.csv(number_additional_points, here("marine/salmon/salmon_farms/output/salmon_points_to_add_final.csv"), row.names = FALSE)
```

@@

write.csv(all_known_farms_final, here("marine/salmon/salmon_farms/output/all_known_salmon_farms_final.csv"), row.names = FALSE)
```

### 2022-01-28 marine/shrimp/shrimp_farms/STEP2_shrimp_farm_locations.Rmd
updated figs and stats with newly run farms
updated for resubmission!

# Indonesia 
Current data is from ASC, only 5 farms, so probably not comprehensive, just a drop in the bucket compared to the produciton of Indonesia. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

@@

Current data is from ASC, only 4 farms, so definitely not comprehensive, just a drop in the bucket compared to the produciton of China. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

@@

write_csv(farms_chn, here("marine/shrimp/shrimp_farms/data/known_farm_sites_2_4_5/farms_chn.csv"))

>>>>

# Indonesia 
Current data is from ASC, only 5 farms, so probably not comprehensive, just a drop in the bucket compared to the production of Indonesia. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

@@

Current data is from https://zenodo.org/record/3881612 and ASC, only 435 + 4 = 439 farms, so probably not comprehensive, just a drop in the bucket compared to the produciton of China. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

@@

farms_chn_2 <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/China/split_spp/farms_shrimp_chn.csv") %>%
  dplyr::select(-id) %>%
  mutate(data_quality = 2)

farms_chn_fin <- rbind(farms_chn, farms_chn_2)

write_csv(farms_chn_fin, here("marine/shrimp/shrimp_farms/data/known_farm_sites_2_4_5/farms_chn.csv"))

### 2022-01-28 marine/united_states/US_mapping.Rmd
updated for resubmission!

**There is comprehensive downloadable data for 5 states: **

 - Maine, Alaska, New Hampshire, New York(?), Connecticut(?)

**There are publicly available map viewers (leases or farm locations) which seem to be comprehensive, but are NOT downloadable for 4 states:**

 - Alabama, North Carolina, Virginia, Rhode Island

**There is general info on lease holders, what they farm, and what municipality or city they are in for 7 states:**

 - California: general areas with acreage for commercial farms
 - Oregon: map of places were commercial harvest/growing is allowed
 - Maryland: Have INDIVIDUAL pdf maps of all lease applications and if they were accepted or not...
 - New Jersey: Have pdf maps of all existing shellfish leases and those issued in 2020
 - Massachusetts: Have number of growers per municipality with acreage and species type; also a subset of farms are available in the NE data viewer.
 - South Carolina: Apparently have 9 oyster farms as of 2021
 - Washington: Approximate locations of commercial net pens.

**Subset of farms downloadable for 1 state:**

 - Florida: A map of mariculture farms which are at risk of oil spills

@@

 - Delaware: Commercial harvest is illegal
 - Georgia: From what I could tell, oyster farming was illegal here as recent as 2020

@@

 - https://www.maine.gov/dmr/aquaculture/leases/aquaculturemap.html
 - I think we will want to use both "limited purpose" and regular leases.
 - Limited purpose leases are for scientific or smaller/shorter term mariculture projects, that last up to 3 years.
 - I think this can be considered "comprehensive". 
 - Update to date as of September 2021.
 - Has species types specified. 
 - Red polygons for the leases. Blue points for the limited purpose. Limited purpose has length and width measurements, so regardless we will have area for both if needed.

@@

@@

Shellfish bed managed

 - https://portal.ct.gov/DOAG/Aquaculture1/Aquaculture/Shellfish-Area-Classifications--Maps
 - Only Shellfish
 - Polygons 
 - I believe that these are large areas that have been **approved** for shellfish aquaculture, but not the exact locations of the aquaculture. So this probably isn't helpful.

@@

ct_shellfish <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Connecticut/Shellfish_Bed_Managed_shp/SHELLFISH_BED_MANAGED.shp"))

ct <- counties("Connecticut", cb = TRUE)

  ggplot() +
  geom_sf(data = ct_shellfish, col = "red") + 
    geom_sf(data = ct, fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Connecticut data", caption = "Not Comprehensive, General 'approved' areas")
```

Another map combined with NY, this is what we need (seaweed and shellfish): https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

 - Lease mapper for Suffolk County: https://gis.suffolkcountyny.gov/portal/apps/View/index.html?appid=6ae708a346f340ada850544fe25ce0fb

 - Shellfish map for NY and CT, this seems like what we would need: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46
 - Pretty easy to pull lat/lon coordinates from the attribute table on the web map.
 - CT has species and culture type. NY only has shellfish on the map

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/ny_shell.png)

 - "Alabama has an emerging oyster industry with 13 farms comprising of 25 acres in 2015. Shellfish farms are a source of healthy protein and can contribute to healthy marine ecosystems by cleaning the water column, providing habitat for other sea life, and helping protect shorelines from storm surges and sea level rise." from https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87
 - This has a map... probably our best source for Alabama!: https://alaquaculture.com/state/#list-of-operating-farms

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/alabama_Farms.png)

@@

 - https://deq.nc.gov/about/divisions/marine-fisheries/licenses-permits-and-leases/shellfish-lease-and-franchise#stats-and-figures
 - Comprehensive for shellfish... just need to figure out how download it. 
 - I believe Shellfish is pretty much the only mariculture that occurs in NC. 
 - Has number totals of leases 
 - Will likely need to send an email to NCDEQ

 ![Has lease info for the entire state](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/nc_map.png)

@@

 - Lease map here: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f
 - No download option.
 - Julia Livermore: julia.livermore@dem.ri.gov

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/rhode_island_map.png)

@@

@@

 - https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87
 - Current as of 2015 
 - Can definitely use for California to cross reference
 - Has partial data for other states too

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/all_states_partial.png)

@@

@@

>>>>

**There is comprehensive downloadable data for 7 states: **

 - Maine, Alaska, New Hampshire, Connecticut, North Carolina, Rhode Island, Alabama

**There are publicly available map viewers (leases or farm locations) which seem to be comprehensive, but are NOT downloadable for 1 state:**

 - Virginia

**There is general info on lease holders, what they farm, and what municipality or city they are in for 6 states:**

 - California: general areas with acreage for commercial farms
 - Oregon: map of places were commercial harvest/growing is allowed
 - New Jersey: Have pdf maps of all existing shellfish leases and those issued in 2020
 - Massachusetts: Have number of growers per municipality with acreage and species type; also a subset of farms are available in the NE data viewer.
 - South Carolina: Apparently have 9 oyster farms as of 2021
 - Washington: Approximate locations of commercial net pens.

**Subset of farms or outdated downloadable for 2 states:**

 - Florida: A map of mariculture farms which are at risk of oil spills
 - Maryland: Have INDIVIDUAL pdf maps of all lease applications and if they were accepted or not... And was able to download lease data from 2013.

@@

 - Delaware: Commercial harvest is illegal; apparently one farm but not production (Halley)
 - Georgia: From what I could tell, oyster farming was illegal here as recent as 2020
 - New York: No production (Halley)

@@

 - Viewer: https://www.maine.gov/dmr/aquaculture/leases/aquaculturemap.html
 - Download: https://dmr-maine.opendata.arcgis.com/datasets/maine::mainedmr-aquaculture-aq-leases/about
 - I think we will want to use regular leases only.
 - Limited purpose leases are for scientific or smaller/shorter term mariculture projects, that last up to 3 years.
 - I think this can be considered "comprehensive". 
 - Update to date as of 2019.
 - Has species types specified.

@@

## leave out limited purpose 
```

@@

@@

Map combined with NY, this is what we need (seaweed and shellfish). We will extract the data using the methods described below: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

Current as of 2019

Extract data following these protocols: https://stackoverflow.com/questions/50161492/how-do-i-scrape-data-from-an-arcgis-online-map/50213419#50213419

CT points json: https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/0/query?where=0%3D0&outFields=%2A&f=json

CT polygons json: https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/1/query?where=0%3D0&outFields=%2A&f=json

Roadmap for polygons

Viewer: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

Item data endpoint (just replace id): https://www.arcgis.com/sharing/rest/content/items/f8799cefbb4c4751a209710d14b9ad46/data

Search for layer you want and grab link ("CT Aquaculture Gear): https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/1

Run query using that link to output geojson: https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/1/query?where=0%3D0&outFields=%2A&f=geojson

And copy-paste the text in here to save as geojson: https://geojson.io

Check the downloaded geojson file: 
```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Connecticut/CT_Aquaculture_gear/CT_aquaculture_gear.geojson",  what = "sp") %>%
  st_as_sf()

mapview(spdf)

# library(geojsonio)
# spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Connecticut/CT_aquacultre_points/CT_aquaculture_points.geojson",  what = "sp") %>%
#   st_as_sf()
# 
# 
# plot(spdf)
# 
# mapview(spdf)

@@

 - https://deq.nc.gov/about/divisions/marine-fisheries/licenses-permits-and-leases/shellfish-lease-and-franchise#stats-and-figures
 - Comprehensive for shellfish... just need to figure out how download it. 
 - I believe Shellfish is pretty much the only mariculture that occurs in NC. 
 - Has number totals of leases.
 - Download using protocol described for CT.
 - Current as of 12/21/2021

Roadmap to download

Viewer: https://www.arcgis.com/apps/webappviewer/index.html?id=de86f3bb9e634005b12f69a8a5947367&extent=-8551979.8781%2C4121555.1994%2C-8515290.1046%2C4140072.0696%2C102100

Item data endpoint (just replace id): https://www.arcgis.com/sharing/rest/content/items/de86f3bb9e634005b12f69a8a5947367/data

Search for layer you want and grab link ("Leases" is the layer you want): https://services2.arcgis.com/kCu40SDxsCGcuUWO/arcgis/rest/services/Lease_DB_View/FeatureServer/2

Run query using that link to output geojson: https://services2.arcgis.com/kCu40SDxsCGcuUWO/arcgis/rest/services/Lease_DB_View/FeatureServer/2/query?where=0%3D0&outFields=%2A&f=geojson

And copy-paste the text in here to save as geojson: https://geojson.io. Once it shows in the map, click "Save" and Save as a geojson file.

Check to see if it worked
```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/North_Carolina/NC_shellfish_leases.geojson",  what = "sp") %>%
  st_as_sf()

mapview(spdf)

 - Lease map here: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f
 - Julia Livermore: julia.livermore@dem.ri.gov
 - Current as of August 2021
 - Shellfish and kelp

Roadmap to download

Viewer: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f

I found my way to here:
https://risegis.ri.gov/hosting/rest/services/RIDEM/Marine_Fisheries_Map_Layers/MapServer/0

Then run query using that link to output geojson: https://risegis.ri.gov/hosting/rest/services/RIDEM/Marine_Fisheries_Map_Layers/MapServer/0/query?where=0%3D0&outFields=%2A&f=geojson

And copy-paste the text in here to save as kml (for some reason, saving as geojson was not working for this one): https://geojson.io. Once it shows in the map, click "Save" and Save as a kml file.

Check to see if it worked

```{r}

rhode_island <- read_sf("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Rhode_Island/ri_aq_sites.kml")

mapview(rhode_island)

 - "Alabama has an emerging oyster industry with 13 farms comprising of 25 acres in 2015. Shellfish farms are a source of healthy protein and can contribute to healthy marine ecosystems by cleaning the water column, providing habitat for other sea life, and helping protect shorelines from storm surges and sea level rise." from https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87
 - This has a map : https://alaquaculture.com/state/#list-of-operating-farms
 - Here is an even better map with Shellfish leases!: https://aldcnr.maps.arcgis.com/apps/webappviewer/index.html?id=a32dad8dacd249ea86bcb80dd951a424

Roadmap to download

Viewer: https://aldcnr.maps.arcgis.com/apps/webappviewer/index.html?id=a32dad8dacd249ea86bcb80dd951a424

Item data endpoint (just replace id): https://www.arcgis.com/sharing/rest/content/items/a32dad8dacd249ea86bcb80dd951a424/data

Search for layer you want and grab link ("Leases" is the layer you want): https://services7.arcgis.com/jcVUvxpWnajErgCr/arcgis/rest/services/Shellfish_Riparian_Leases_Revised/FeatureServer/0

Run query using that link to output geojson: https://services7.arcgis.com/jcVUvxpWnajErgCr/arcgis/rest/services/Shellfish_Riparian_Leases_Revised/FeatureServer/0/query?where=0%3D0&outFields=%2A&f=geojson

And copy-paste the text in here to save as geojson: https://geojson.io. Once it shows in the map, click "Save" and Save as a geojson file.

Leases current as of: 11/29/2021

```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Alabama/Alabama_Shellfish_Leases.geojson",  what = "sp") %>%
  st_as_sf()

mapview(spdf)

@@

@@

List of farms for VA and MD (addresses are on land though): https://oysterrecovery.org/oyster-farms/

Maryland has a viewer for 2013 active leases: https://maryland.maps.arcgis.com/apps/webappviewer/index.html?id=f7db2acc81554601b44c13733a66a60f

 - This viewer has a layer named "ActiveOysterLeases", however, it looks like the last time it was updated is 2013.. I will still download this. Worst case we use this.

I found the layer here: https://gisapps.dnr.state.md.us/arcgis2/rest/services/Biota/ActiveOysterLeases/MapServer/0

Run query using that link to output json (for some reason it won't let me output geojson): https://gisapps.dnr.state.md.us/arcgis2/rest/services/Biota/ActiveOysterLeases/MapServer/0/query?where=0%3D0&outFields=%2A&f=json

```{r include = FALSE}

# test out package to download
devtools::install_github("yonghah/esri2sf")
library("esri2sf")

url <- "https://gisapps.dnr.state.md.us/arcgis2/rest/services/Biota/ActiveOysterLeases/MapServer/0"
df <- esri2sf(url)
mapview(df)

#st_write(df, file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maryland/maryland_2013_oyster_leases.shp"))

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
md <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maryland/maryland_2013_oyster_leases.shp"))
```

@@

 - https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87
 - another link to map? https://aop.maps.arcgis.com/apps/Shortlist/index.html?appid=32864a80775b4f8abaeb072f0718ee57
 - Current as of 2015 
 - Can definitely use for California to cross reference
 - Has partial data for other states too
 - ask Halley for contact

@@

 - According to Halley, there is currently no production in New York state, especially since their coastline is so small.

@@

## Climate stressors for marine aquaculture

### 2022-01-28 workflow/calc_dist_to_port.Rmd
cleaned up scripts and READMEs
updated for resubmission!

@@

## ok so that works... i just to calculate the distance from each point in the whole known location dataset, and find the min of each column

@@

write.csv(dist_ports, "int/distance_to_ports_raw.csv")

@@

mean(spp_group_dist$closest_port_km) # 31.30463 km
sd(spp_group_dist$closest_port_km) # 32.333 km
range(spp_group_dist$closest_port_km) #  0.1312871 to 531.2067646
median(spp_group_dist$closest_port_km) # 18.85009 km

@@

spp_group_dist_final <- spp_group_dist %>%
  group_by(type) %>%
  summarise(dist_port = mean(closest_port)) %>%
  ungroup()

>>>>

This script calculates an average distance to closest port from all of the known mariculture locations.

@@

## ok so that works... i just need to calculate the distance from each point in the whole known location dataset, and find the min of each column

@@

# write.csv(dist_ports, "int/distance_to_ports_raw.csv")

@@

mean(spp_group_dist$closest_port_km) # 31.30463 km v1; 39.82486 v2; 39.65138 v3 scotland fix
sd(spp_group_dist$closest_port_km) # 32.333 km v1; 34.9805 v2; 34.94368 v3
range(spp_group_dist$closest_port_km) #  0.1312871 to 531.2067646 v1; 0.1309383 533.2698526 v2; 0.1309383 533.2698526 v3
median(spp_group_dist$closest_port_km) # 18.85009 km v1; 28.56732 v2; 28.33052 v3

@@

spp_group_dist_final <- spp_group_dist %>%
  group_by(type) %>%
  summarise(dist_port = mean(closest_port_km)) %>%
  ungroup()
